Making multiple-locus association mapping on a genome-wide scale routine

### 2017-11-08 AM+draft.tex
8nov17 night

BigRR, FaST-LASSO, and glmnet are all based on regularization methods. Regression based regularization methods 
are well suited to solving problems where the number of predictors (p) is far greater than the number of samples (n).  It 
should come as no surprise then that multiple locus association mapping are dominated by regularization-based approaches.
Regularisation methods are able to yield solutions to problems that OLS cannot. They introduce a penalty function that balances
bias against reduced variance.  BigRR, an R package, implements gen

>>>>

Of the five multiple locus methods considered, BigRR, LMM-Lasso, and glmnet,  implement regularization methods for parameter
estimation. Regularization methods are well suited to problems where the number of predictors is far greater than the 
number of samples. It is not surprising then that regularization methods have attracted considerable interest, primarily for 
genomic prediction (refs) but also association mapping (refs). BigRR is an R package that implements generalized ridge regression
which is an improvement on standard ridge regression. In standard ridge regression, coefficients of the predictors are shrunk 
to zero by a constant amount.  In generalized ridgre regression, the amount of shrinkage is allowed to vary, giving opportunity for 
a more realistic model to be specified. The model used in BigRR is 
$y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$.

BigRR, LMM-LASSO, and glmnet are all based on regularization methods. Regression based regularization methods 
are well suited to solving problems where the number of predictors (p) is far greater than the number of samples (n).  It 
is not surprising that multiple locus association mapping are dominated by regularization-based approaches. Regularisation-based 
approaches model the effect of all the snp on a trait simultaneously. 
Regularisation methods are able to yield solutions to problems that OLS cannot. They introduce a penalty function that balances
bias against reduced variance.  BigRR, an R package, implements generalised ridge regression. Here, the underlying model 
can be formulated as a linear mixed model that is
 $y = X \beta + M u + e$ where $u^{(L \times 1)}$ is a vector of snp effects, and each snp 
effect is normally distributed with its own variance such that $u_l \sim N(0, \sigma^2_l); l=1,2, \ldots, L$. The size of the random 
snp effect is a measure of the strength of association between the snp and trait. The parameters are estimated via an
algorithm that makes clever use of the equivalence between the above model and animal models \ref{}.

Lmm-Lasso  \cite{rakitsch2013lasso}, a stand-alone computer program, implements an estimation 
procedure based on lasso for the (full) model 
$y =  \sum_{l=1}^L m_l \theta_l +   g  + e$ where $g^{(n ; \times 1)}$ is a vector of random polygenic effects such that 
$g \sim N(0, \sigma^2_g K)$, and $K^{(n ; \times \; n)}$ is a realised relationship matrix reconstructed from the snp information. 
Lasso is popular due to its capacity to perform parameter estimation and variable selection, simultaneously. Lasso though is demanding, computationally. To reduce this problem, a two-step 
estimation procedure is adopted.  First, the above model without any snp effects is fitted. Second, the unknown variance components in the full model are replaced by their estimates that were calculated in the first step.  This turns the model into a simpler linear model for which lasso is far more tractable, computationally.

Glmnet, a general purpose R package, uses elastic net for parameter estimation.

makes use of elastic net. 
The model used for analysis is 
$y = X \beta + \sum_{l=1}^L m_l \theta_l + e$ where $\theta_l; l=(1,2, \ldots, L)$ is the coefficient for the $l$th snp. 
Glmnet implements elastic net for the estimation of the parameters. Elastic net is a regularisation method where the 
penalty function is a mixture of 
$\ell_1$-norm (lasso) and $\ell_2$-norm (ridge regression) penalties  \cite{zou2005regularization}. 
This allows elastic net to avoid the limitations of lasso 
while still being able to perform shrinkage and variable selection, simultaneously \cite{zou2005regularization}.

Here, 
the 
amount of shrinkage is allowed to vary across the predictors (snp effects) where in standard ridge regression, the amount of 
shrinkage is constant. FaST-LMM, a stand-alone computer program, implements a two-step lasso approach.
